Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} model system

### 2014-04-25 paper_enahpc14/Sec4:Experimental-results.tex
Manuel's changes

We  approach  the  assessment  of  the energy  footprint  and  overall
performance   of   \cosmoart   with  two   important  metrics:
\textit{time-to-solution} (TTS) and \textit{energy-to-solution} (ETS).
TTS refers to  the total wall clock time  of the application execution
time. ETS  is the amount of  energy spent to  achieve results.  Energy
consumption  is  assessed by  sampling  the  instantaneous peak  power
during execution which  is then averaged and multiplied  by the TTS to
determine  ETS.   Whenever  possible,  multiple production  runs  were
performed  to  illustrate the  reproducibility  of  the baseline,  and
quantify the  significant uncertainties  in the power  measurement, as
dictated by the available technology.

@@

\begin{figure}[htbf]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{Figs/NRJ_benchmark_Monch.eps}
    \caption{\textsc{Monch}: Isola E1 Rack 2 Total Power.}
    \label{fig:1}
  \end{center}
\end{figure}

\begin{figure}[htbf]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{Figs/NRJ_benchmark_Pilatus.eps}
    \caption{\textsc{Pilatus}: Isola HD Total Power.}
    \label{fig:2}
  \end{center}
\end{figure}

Figures~\ref{fig:1}   and   \ref{fig:2}   account   respectively   for
\textsc{Monch}'s Isola E1 Rack  2 and \textsc{Pilatus}' Isola HD total
power measurements for  1-day or 2-days simulations. On  the Intel Ivy
Bridge EP  based cluster (i.e.  \textsc{Monch}),  the 1-day simulation
was issued  only twice due  to usage restrictions. As  time resolution
was set to one update every  5 minutes for power sampling, the average
power consumption was computed by considering 6 values for each single
run.  On the Intel Xeon E5 based cluster (i.e.  \textsc{Pilatus}), the
1-day simulation  was issued  four times and  a 2-days run  only once.
Similarly, the average power consumption was computed by considering 4
values  for  each  single  1-day  run  and 9  values  for  the  2-days
run. Corresponding results are gathered in Table~\ref{tab:3}.

@@

The 24 hours simulation time-to-solution and energy-to-solution 
results for Tintorrum are shown in Figure~\ref{fig:4}.  As expected, the mean energy
consumed by COSMO-ART when using the MPI blocking mode is less than that
consumed by the polling mode. This is due to the blocking-waits performed by the cores
while using the blocking mode, in contrast of the polling-waits that take place when
polling mode is used. From the point of view of time, the results are quite
surprising since the blocking mode was expected to take less time than the polling
one.

\begin{figure}[htbf]
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-0.eps}
  \caption{Time-to-solution and energy-to-solution comparison between
    Xeon E5 and Ivy Bridge-EP architectures for a 24h simulation}
  \label{fig:3}
\end{figure}

\begin{figure}[htbf]
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-5.eps}
  \caption{Mean time-to-solution and energy-to-solution on
    \textsc{Tintorrum} for a 24h simulation - MPI Polling and Blocking
    modes}
  \label{fig:4}
\end{figure}

\subsection{Power-performance tracing of \cosmoart}
\label{subsec:4.3}

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution. First we run \cosmoart for a full-day simulation
and analyze the power consumption. Second we correlate performance with
power traces of two 1-hor time frames during the day: mid-day (from 11h-12h) and
midnight (23-24h).

@@

In order to obtain an initial overview of the power dissipation of the system model, Figure~\ref{fig:9} depicts the power profiles of a full-day simulation of \cosmoart leveraging the  polling and blocking MPI policies on the top and bottom plots, respectively. The power profiles shown account for the sum of dissipated power of the 16 nodes used in \tinto, captured with our PDUs. The first observation for the polling MPI policy is the quite plain power profile it generates. However, the first and last hours of the day along a 24h simulation, dissipate slightly less power than the hours than the mid-day hours. We relate these variations due to the chemical reactions and aerosols, computed by ART, that are taking place during the daylight hours but not during night. In this case peak power dissipation is about 3821\,W.
Looking at the power profile using the blocking MPI policy, we can easily observe
that the pattern is repeated favoring the reduction of power during nighttime hours in a more noticeable way. The peak power dissipation in this mode is 3768\,W, which means a reduction of 50\,W in total or, in other words, 1,\% of reductuon in the peak power. Also, we observe power drops in each hour simulated, basically related to periods in which the processes were performing blocking-waits instead of polling-waits, thus yielding the CPU to a low usage during this phases. Our final observation is the reduction in execution time when the blocking MPI policy is leveraged. One can expect that the costs of the interruptions used in the blocking can lead to an increase of the total execution time, however, applications with extensive computing may show better perforamnce with the blocking mode. This plays in favour for our results since the energy consumed is also less by just changing the behavior of
the OpenMPI library from the polling to the blocking policy.

>>>>

We  approach  the  assessment  of  the energy  footprint  and  overall
performance   of   \cosmoart   with  two   important  metrics:
\textit{time-to-solution} (TTS) and \textit{energy-to-solution} (ETS).
TTS refers to  the total wall clock time  of the application execution
time. ETS  is the amount of  energy spent to  achieve results.  Energy
consumption  is  assessed by  sampling  the  power during execution which is then averaged and multiplied  by the TTS to determine  ETS. Whenever  possible,  multiple production  runs  were
performed  to  illustrate the  reproducibility  of  the baseline,  and
quantify the  significant uncertainties  in the power  measurement, as
dictated by the available technology.

@@

\begin{figure}[htbf]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{Figs/NRJ_benchmark_Monch.eps}
    \caption{\monch: Isola E1 Rack 2 Total Power.}
    \label{fig:1}
  \end{center}
\end{figure}

\begin{figure}[htbf]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{Figs/NRJ_benchmark_Pilatus.eps}
    \caption{\pilat: Isola HD Total Power.}
    \label{fig:2}
  \end{center}
\end{figure}

Figures~\ref{fig:1}   and   \ref{fig:2}   account   respectively   for
\monch's Isola E1 Rack  2 and \pilat' Isola HD total
power measurements for  1-day or 2-days simulations. On  the Intel Ivy
Bridge EP  based cluster (i.e.  \monch),  the 1-day simulation
was issued  only twice due  to usage restrictions. As  time resolution
was set to one update every  5 minutes for power sampling, the average
power consumption was computed by considering 6 values for each single
run.  On the Intel Xeon E5 based cluster (i.e.  \pilat), the
1-day simulation  was issued  four times and  a 2-days run  only once.
Similarly, the average power consumption was computed by considering 4
values  for  each  single  1-day  run  and 9  values  for  the  2-days
run. Corresponding results are gathered in Table~\ref{tab:3}.

@@

The TTS and ETS results for a full-day simulation on \tinto are shown in Figure~\ref{fig:4}.  As expected, the mean energy consumed by COSMO-ART when using the blocking MPI policy is about 2\,W less than that consumed by the polling mode. This is due to the blocking-waits performed by the CPU cores while using the blocking mode, in contrast of the polling-waits that take place when polling mode is used. From the point of view of the execution time, there is a slight reduction when the blocking MPI policy is leveraged. One can expect that the costs of the interruptions used in the blocking MPI policy can lead to an increase of the total execution time, however, applications with extensive computing may show better perforamnce with the blocking modes. This plays in favour for our results since the energy consumed is also less by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

\begin{figure}[htbf]
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-0.eps}
  \caption{Time-to-solution and energy-to-solution comparison between
    Xeon E5 and Ivy Bridge-EP architectures for a 24h simulation.}
  \label{fig:3}
\end{figure}

\begin{figure}[htbf]
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-5.eps}
  \caption{Mean time-to-solution and energy-to-solution on
    \textsc{Tintorrum} for a 24h simulation using the polling and blocking MPI policies.}
  \label{fig:4}
\end{figure}

\subsection{Power-performance tracing on \tintorrum}
\label{subsec:4.3}

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution. First we run \cosmoart for a full-day simulation
and analyze the power consumption. Second we correlate performance with
power traces of two 1-hor time frames during the day: mid-day (from 11h-12h) and
midnight (23-24h). All the experiments were performed using 192 \cosmoart processes
mapping 12 processes in each of node.

@@

In order to obtain an initial overview of the power dissipation of the system model, we have obtained the power profiles using our \pmlib framework of a full-day simulation of \cosmoart leveraging the polling and blocking MPI policies, these results are depicted in the top/bottom plots of Figure~\ref{fig:9}, respectvely. o, captured with our PDUs. The first observation for the polling MPI policy is the quite plain power profile it generates. However, the first and last hours of the full-day simulation dissipate slightly less power than the mid-day hours. We relate these variations due to the chemical reactions and aerosols, computed by ART, that are taking place during the daylight hours but not during night. In this case peak power dissipation is about 3821\,W.
Looking at the power profile using the blocking MPI policy, we can easily observe
that the pattern is repeated favoring the reduction of power during nighttime hours in a more noticeable way. The peak power dissipation in this mode is 3768\,W, which means a reduction of 50\,W in total or, in other words, 1,\% of reductuon in the peak power. Also, we observe power drops in each hour simulated, basically related to periods in which the processes were performing blocking-waits instead of polling-waits, thus yielding the CPU to a low usage during this phases. Our final observation is the reduction in execution time when the blocking MPI policy is leveraged in contrast with the polling MPI. One can expect that the costs of the interruptions used in the blocking can lead to an increase of the total execution time, however, applications with extensive computing may show better perforamnce with the blocking mode. This plays in favour for our results since the energy consumed is also less by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

To gain insights about
